---
title: "PHSL4005 Stats. Exam"
author: "Vishaylin J. Mahadeo"
date: "August 31, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      cache = TRUE,
                      fig.align = 'center',
                      fig.height = 8,
                      fig.width = 8,
                      fig.ext = 'png',
                      fig.path = 'figure/')
library(dplyr)
library(readr)
library(tidyr)
```

# Question 1
## Body Temperature between males and female heart rates

```{r Q1_load}
#load data set
df_Q1 <- read_csv('question1.csv')

#view first 6 lines of data set
head(df_Q1, 6)

#tidy data (Collect males)
Q1_male <- df_Q1 %>% 
  filter(male != 'NA') %>%
  select(body_temperature, male)
Q1_male

#tidy data (Collect females)
Q1_female <- df_Q1 %>%
  filter(female != 'NA') %>%
  select(body_temperature, female)
Q1_female

#plot data (males)
male_plot <- plot(x = Q1_male$male, y = Q1_male$body_temperature,
                  main = 'Scatter plot showing male heart rate vs. body temperature',
                  xlab = 'Male heart rate (BPM)',
                  ylab = 'Body temperature (degrees Celcius)')

#linear regression for males
male_rgl <- lm(body_temperature~male, data = Q1_male)
abline(male_rgl)

#print regression line for males
summary(male_rgl)

#check homoskedasticity
male_homosk <- plot(x = male_rgl$fitted, y = male_rgl$residuals,
                    main = 'Homoskedasticity check for males',
                    xlab = 'Fitted', ylab = 'Residuals')
abline(h=0)

#Check Gaussian distribution
qqnorm(male_rgl$residuals)
qqline(male_rgl$residuals)

#plot data (females)
female_plot <- plot(x = Q1_female$female, y = Q1_female$body_temperature,
                    main = 'Scatter plot showing female heart rate vs. body temperature',
                    xlab = 'Female heart rate (BPM)',
                    ylab = 'Body temperature (degrees Celcius)')

#linear regression for females
female_rgl <- lm(body_temperature~female, data = Q1_female)
abline(female_rgl)

#print regression line for females
summary(female_rgl)

#check homoskedasticity
female_homosk <- plot(x = female_rgl$fitted, y = female_rgl$residuals,
                    main = 'Homoskedasticity check for females',
                    xlab = 'Fitted', ylab = 'Residuals')
abline(h=0)

#Check Gaussian distribution
qqnorm(female_rgl$residuals)
qqline(female_rgl$residuals)

```
**Null Hypothesis:** Heart rates in males and females have no association with body temperature.

**Alternative Hypothesis:** Heart rates in males and females have an association with body temperature.

**Assumptions:**

1. a = 0.05

2. Paired, continuous data set.

3. Parametric data set because normal distribution indicated by a normal Q-Q plot.

4. Residuals are normally distributed as shown by homoskedasticity check.

5. Reject null hypothesis if p < a.

6. Pearson's correlation will be used to test hypothesis.

```{r Q1_test}
#male Pearson's correlation
male_cor <- with(Q1_male,
                 cor.test(x = male, y = body_temperature,
                          method = 'pearson'))
male_cor

#male paired t-test
male_test <- t.test(Q1_male$male, Q1_male$body_temperature, paired = TRUE)
male_test

#female Pearson's correlation
female_cor <- with(Q1_female,
                 cor.test(x = female, y = body_temperature,
                          method = 'pearson'))
female_cor

#female paired t-test
female_test <- t.test(Q1_female$female, Q1_female$body_temperature, paired = TRUE)
female_test
```
**Results:**

Males presented with t(64) = 50.81; p < 0.001 and R^2^= 0.188; F(0.384)=2.313

Females presented with t(64) = 37.57; p < 0.001 and R^2^= 0.284; F(0.399)=5.528

p < a, thus reject null hypotheis and accept alternative.

**Conclusion:** In conclusion there is a weak, positive association between heart rates in both males and females and their body temperature.

****

# Question 2
## Handedness compared to ataxic walking and intoxication

```{r Q2_load}
#load data set
df_Q2 <- read.csv('question2.csv')

#select data
Q2_data <- df_Q2 %>%
  select(handedness, final_position)

Q2_data2 <- df_Q2 %>%
  select(handedness, first_stumble)
         
#display data
head(Q2_data, 6)
head(Q2_data2)

#Counts (handedness and first stumble)
rightS_side <- Q2_data2 %>%
  filter(handedness == 1) %>%
filter(first_stumble == 1)
  
rightS_count <- count(rightS_side)
rightS_count

leftS_side <- Q2_data2 %>%
  filter(handedness == 0) %>%
  filter(first_stumble == 0)
  
leftS_count <- count(leftS_side)
leftS_count

rightS_left <- Q2_data2 %>%
  filter(handedness == 1) %>%
  filter(first_stumble == 0)

rightS_left_count <- count(rightS_left)
rightS_left_count

leftS_right <- Q2_data2 %>%
  filter(handedness == 0) %>%
  filter(first_stumble == 1)

left_rightS_count <- count(leftS_right)
left_rightS_count


#Counts (handedness and position and end of walk)
right_side <- Q2_data %>%
  filter(handedness == 1) %>%
filter(final_position == 1)
  
right_count <- count(right_side)
right_count

left_side <- Q2_data %>%
  filter(handedness == 0) %>%
  filter(final_position == 0)
  
left_count <- count(left_side)
left_count

right_left <- Q2_data %>%
  filter(handedness == 1) %>%
  filter(final_position == 0)

right_left_count <- count(right_left)
right_left_count

left_right <- Q2_data %>%
  filter(handedness == 0) %>%
  filter(final_position == 1)

left_right_count <- count(left_right)
left_right_count

#Form table
Q2_dataFrame <- data.frame(Handedness = c('Right', 'Left'),
                           Direction_first_stumble = c(53,22,22,54),
                           Position_at_end = c(48,27,36,40))
Q2_dataFrame
```

**Null Hypothesis:** Handedness is not associated with ataxic walking in intoxicated individuals.

**Alternative Hypothesis:** Handedness is associated with ataxic walking in intoxicated individuals.

**Assumptions:**

1. a = 0.05

2. Paired data set

3. Parametric data set

4. Reject null hypothesis if p < a

5. Fisher's exact test with a Holm adustment method will be required to test hypothesis.

```{r Q2_test}
#
```
****

# Question 3
## Can running time be used to predict calories burned

```{r Q3_load}
#load data set
df_Q3 <- read_csv('question3.csv')

#plot data set
Q3_plot <- plot(time~calories, data = df_Q3,
                main = 'Scatter plot representing running time vs. calories burnt' ,
                xlab = 'Running Time (s)', ylab = 'Calories consumed(cal)')

#tidy data
Q3_run <- select(df_Q3, time, calories)

#linear regression 
run_rgl <- lm(calories~time, data = Q3_run)
abline(run_rgl)

#print regression line for males
summary(run_rgl)

#check homoskedasticity
run_homosk <- plot(x = run_rgl$fitted, y = run_rgl$residuals,
                    main = 'Homoskedasticity check for running time vs. calories consumed',
                    xlab = 'Fitted', ylab = 'Residuals')
abline(h=0)

#Check Gaussian distribution
qqnorm(run_rgl$residuals)
qqline(run_rgl$residuals)
```

**Null Hypothesis:** Running time (s) cannot be used to predict number of calories consumed (cal) during a run.

**Alternative Hypothesis:** Running time (s) can be used to predict number of calories consumed (cal) during a run.

**Assumptions:**

1. a = 0.05

2. Paired, interval data set with no outliers.

3. Gaussian residual distribution(i.e.normal distribution) indicated by normal Q-Q plot. 

4. Parametric data set.

5. Residuals are heteroskedastic because they show spread around the abline at 0.

6. Running time is measured without error.

7. Observations are independent.

8. If p < a reject null hypothesis.

9. A Pearson's correlation will be used to test hypothesis.

```{r Q3_test}
#Pearson's correlation
run_cor <- with(Q3_run,
                 cor.test(x = time, y = calories,
                          method = 'pearson'))
run_cor

#calculate regression coefficient
regco <- coefficients(run_rgl)
regco

#fit coeffiecients into formula
origTime <- 30
decTimeSec <- origTime * 60
run_time <- mean(df_Q3$run) - decTimeSec
CaloriesC <- regco[1] + regco[2] * run_time
CaloriesC
```
**Results:** 

t(17) = -3.771; p = 0.002 and R^2^= -0.674; F(17)= 14.22

p < a, thus reject null hypothesis.

**Conclusion:** Running time has a strong, negative correlation with calories burnt. Therefore the shorter runs burn more calories than longer runs.

Furthermore if she decreased her running time by 30 minutes she would burn 1441.128 calories.

****

# Questions 4
## The relationship between project mark and final mark for BSc Honours Physiology students over 10 years.

```{r Q4_dirty, fig.ext= 'pdf'}
foo <- rnorm(10000, mean = 60, sd = 3) # final mark
bar <- rnorm(10000, mean = 68, sd = 5) # project mark
baz <- rep(seq(from = 1997, to = 2006), each = 1) # years

year <- sample(baz, 150, replace = TRUE,
               prob = c(0.05, 0.05, 0.08, 0.08, 
                          0.1, 0.13, 0.14, 0.13, 0.12, 0.12))
project_mark <- sample(bar, 150, replace = TRUE)
final_mark <- sample(foo, 150, replace = TRUE)

plot_data <- data_frame(year = year,
                        project_mark = round(project_mark, 1),
                        final_mark = round(final_mark, 1)) %>%
    arrange(year)
#Print out data
head(plot_data, 6)

#Scatter plot of data
mark_plot <- plot(x = project_mark, y = final_mark,
     main = 'Scatter plot representing project mark vs. final mark of BSc Honours Physiology students',
     xlab = 'Project mark (%)', ylab = 'Final year mark (%)')

mark_regl <- lm(final_mark~project_mark)
abline(mark_regl)
summary(mark_regl)
```
**Explanation of axis:**

**x-axis =** project mark (independent variable) because is not affected by final year mark, however affects final year mark.

**y-axis =** final year marl (dependent variable) because is affected by individual project marks over the years.

****
