---
title: "PHSL4005 Stats. Exam"
author: "Vishaylin J. Mahadeo"
date: "August 31, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      cache = TRUE,
                      fig.align = 'center',
                      fig.height = 8,
                      fig.width = 8,
                      fig.ext = 'png',
                      fig.path = 'figure/')
library(dplyr)
library(readr)
library(tidyr)
```

# Question 1
## Body Temperature between males and female heart rates

```{r Q1_load}
#load data set
df_Q1 <- read_csv('question1.csv')

#view first 6 lines of data set
head(df_Q1, 6)

#tidy data (Collect males)
Q1_male <- df_Q1 %>% 
  filter(male != 'NA') %>%
  select(body_temperature, male)
Q1_male

#tidy data (Collect females)
Q1_female <- df_Q1 %>%
  filter(female != 'NA') %>%
  select(body_temperature, female)
Q1_female

#plot data (males)
male_plot <- plot(x = Q1_male$male, y = Q1_male$body_temperature,
                  main = 'Scatter plot showing male heart rate vs. body temperature',
                  xlab = 'Male heart rate (BPM)',
                  ylab = 'Body temperature (degrees Celcius)')

#linear regression for males
male_rgl <- lm(body_temperature~male, data = Q1_male)
abline(male_rgl)

#print regression line for males
summary(male_rgl)

#check homoskedasticity
male_homosk <- plot(x = male_rgl$fitted, y = male_rgl$residuals,
                    main = 'Homoskedasticity check for males',
                    xlab = 'Fitted', ylab = 'Residuals')
abline(h=0)

#Check Gaussian distribution
qqnorm(male_rgl$residuals)
qqline(male_rgl$residuals)

#plot data (females)
female_plot <- plot(x = Q1_female$female, y = Q1_female$body_temperature,
                    main = 'Scatter plot showing female heart rate vs. body temperature',
                    xlab = 'Female heart rate (BPM)',
                    ylab = 'Body temperature (degrees Celcius)')

#linear regression for females
female_rgl <- lm(body_temperature~female, data = Q1_female)
abline(female_rgl)

#print regression line for females
summary(female_rgl)

#check homoskedasticity
female_homosk <- plot(x = female_rgl$fitted, y = female_rgl$residuals,
                    main = 'Homoskedasticity check for females',
                    xlab = 'Fitted', ylab = 'Residuals')
abline(h=0)

#Check Gaussian distribution
qqnorm(female_rgl$residuals)
qqline(female_rgl$residuals)

```
**Null Hypothesis:** Heart rates in males and females have no association with body temperature.

**Alternative Hypothesis:** Heart rates in males and females have an association with body temperature.

**Assumptions:**

1. a = 0.05

2. Paired, continuous data set.

3. Parametric data set because normal distribution indicated by a normal Q-Q plot.

4. Residuals are normally distributed as shown by homoskedasticity check.

5. Reject null hypothesis if p < a.

6. Pearson's correlation will be used to test hypothesis.

```{r Q1_test}
#male Pearson's correlation
male_cor <- with(Q1_male,
                 cor.test(x = male, y = body_temperature,
                          method = 'pearson'))
male_cor

#male Pearson's correlation
female_cor <- with(Q1_female,
                 cor.test(x = female, y = body_temperature,
                          method = 'pearson'))
female_cor
```
**Results:**

Males presented with t(63) = 1.521; p = 0.1333 and R^2^ (-0.058 to 0.413)= 0.188.

Females presented with t(63) = 2.351; p = 0.022 and R^2^ (0.043 to 0.494)= 0.284.

p < a, thus reject null hypotheis and accept alternative.

**Conclusion:** In conclusion there is a weak, positive association between heart rates in males and females and their body temperature.

****

# Question 2
## Handedness compared to ataxic walking and intoxication

```{r Q2_load}
#load data set
df_Q2 <- read.csv('question2.csv')

#plot data set

```

**Null Hypothesis:** 

**Alternative Hypothesis:**

**Assumptions:**

1.

2.

3.

4.

5.

```{r Q2_test}

```
****

# Question 3
## Can running time be used to predict calories burned

```{r Q3_load}
#load data set
df_Q3 <- read_csv('question3.csv')

#plot data set
Q3_plot <- plot(time~calories, data = df_Q3,
                main = 'Scatter plot representing running time vs. calories burnt' ,
                xlab = 'Running Time (s)', ylab = 'Calories consumed(cal)')

#tidy data
Q3_run <- select(df_Q3, time, calories)

#linear regression 
run_rgl <- lm(calories~time, data = Q3_run)
abline(run_rgl)

#print regression line for males
summary(run_rgl)

#check homoskedasticity
run_homosk <- plot(x = run_rgl$fitted, y = run_rgl$residuals,
                    main = 'Homoskedasticity check for running time vs. calories consumed',
                    xlab = 'Fitted', ylab = 'Residuals')
abline(h=0)

#Check Gaussian distribution
qqnorm(run_rgl$residuals)
qqline(run_rgl$residuals)
```

**Null Hypothesis:** Running time (s) cannot be used to predict number of calories consumed (cal) during a run.

**Alternative Hypothesis:** Running time (s) can be used to predict number of calories consumed (cal) during a run.

**Assumptions:**

1. a = 0.05

2. Paired, interval data set with no outliers.

3. Gaussian residual distribution(i.e.normal distribution) indicated by normal Q-Q plot. 

4. Parametric data set.

5. Residuals are heteroskedastic because they show spread around the abline at 0.

6. Running time is measured without error.

7. Observations are independent.

8. If p < a reject null hypothesis.

9. A Pearson's correlation will be used to test hypothesis.

```{r Q3_test}
#Pearson's correlation
run_cor <- with(Q3_run,
                 cor.test(x = time, y = calories,
                          method = 'pearson'))
run_cor

#calculate regression coefficient
regco <- coefficients(run_rgl)
regco

#fit coeffiecients into formula
origTime <- 30
decTimeSec <- origTime * 60
run_time <- mean(df_Q3$run) - decTimeSec
CaloriesC <- regco[1] + regco[2] * run_time
CaloriesC
```
**Results:** 

t(17) = -3.771; p = 0.002 and R^2^(-0.864 to -0.318) = -0.674.

p < a, thus reject null hypothesis.

**Conclusion:** Running time has a strong, negative correlation with calories burnt. Therefore the shorter runs burn more calories than longer runs.

****

# Questions 4
## 



****
